---
title: "Analysis of Boston Properties"
author: "Dawn Graham, Sylvie Xiang"
output:
  html_document:
    fig_width: 8
    code_folding: hide
    highlight: tango
    toc: true
    toc_float:
      collapsed: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


# The Data Set

# Preparing the data
- Import the data set into R.
- Document the steps for the import process and any preprocessing had
to be done prior to or after the import. Any R code used in the process
should be included.

## 2021 Boston Property Assessment Data
```{r echo=TRUE}
library(plotly)
library(plyr)

pa <- read.csv("https://data.boston.gov/dataset/e02c44d2-3c64-459c-8fe2-e1ce5f38a035/resource/c4b7331e-e213-45a5-adda-052e4dd31d41/download/data2021-full.csv",
               colClasses=c("ZIPCODE"="character"))

# Get only desired columns
cols <- c("PID", "CITY", "ZIPCODE", "LU_DESC", "OWN_OCC", "LIVING_AREA", "TOTAL_VALUE", "YR_BUILT", "EXT_COND", "BED_RMS", "FULL_BTH")

pa <- pa[, cols]

# Get only "SINGLE FAM DWELLING"
pa <- pa[pa$LU_DESC=="SINGLE FAM DWELLING",]

# Convert string "$ddd,ddd.dd" to numeric
pa$TOTAL_VALUE <- as.numeric(gsub('[$,]', '', pa$TOTAL_VALUE))

# Map values for OWN_OCC
pa$OWN_OCC <- mapvalues(pa$OWN_OCC,
                        from=c("Y", "N"),
                        to=c("Yes", "No"))

summary(pa)
```

# Analyzing the data

## Analysis
- Do the analysis as in Module3 for at least one categorical variable and at least one numerical variable. Show appropriate plots for your data.

### Owner Occupied
```{r echo=TRUE}
# Categorical variable: OWN_OCC (Dawn)
own_occ <- table(pa$OWN_OCC)

prop.table(own_occ)*100

plot_ly(x = names(own_occ),
        y = as.numeric(own_occ),
        type = "bar"
        ) %>%
  layout(title = "Owner Occupied",
         xaxis = list(categoryorder = "category descending"),
         yaxis = list(title = "Single Family Dwellings")
         )
```

### Full Baths
```{r echo=TRUE}
# Numerical variable: FULL_BTH (Dawn)
full_bth <- table(pa$FULL_BTH)
full_bth_names <- as.numeric(names(full_bth))

plot_ly(x = full_bth_names,
        y = as.numeric(full_bth),
        type = "bar"
        )%>%
  layout(title = "Full Bathrooms",
         xaxis = list(title = "Bathrooms",
                      tickvals = seq(1:max(full_bth_names))),
         yaxis = list(title = "Single Family Dwellings")
  )
```
- Do the analysis as in Module3 for at least one set of two or more variables. Show appropriate plots for your data.

## Distribution
- Pick one variable with numerical data and examine the distribution of the data.

## Central Limit Theorem
- Draw various random samples of the data and show the applicability of the
Central Limit Theorem for this variable.

## Sampling
- Show how various sampling methods can be used on your data. What are your
conclusions if these samples are used instead of the whole dataset.
- Implementation of any feature(s) not mentioned in the above specification.