---
title: "Analysis of Boston Properties"
author: "Dawn Graham, Sylvie Xiang"
output:
  html_document:
    fig_width: 8
    code_folding: hide
    highlight: tango
    toc: true
    toc_float:
      collapsed: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.width=8, fig.height=4,
                      echo=FALSE, warning=FALSE, message=FALSE)
```


# The Data Set

# Preparing the data
- Import the data set into R.
- Document the steps for the import process and any preprocessing had
to be done prior to or after the import. Any R code used in the process
should be included.

## 2021 Data
```{r echo=TRUE}
pa21 <- read.csv("data2021-full.csv")

# Get only desired columns
cols <- c("PID", "CITY", "LUC", "OWN_OCC", "LIVING_AREA", "TOTAL_VALUE", "YR_BUILT", "OVERALL_COND", "BED_RMS", "FULL_BTH")

pa21 <- pa21[, cols]

# Get only "SINGLE FAM DWELLING"
pa21 <- pa21[pa21$LUC==101,]

# Convert string "$ddd,ddd.dd" to numeric
pa21$TOTAL_VALUE <- as.numeric(gsub('[$,]', '', pa21$TOTAL_VALUE))

summary(pa21)
```

## 2019 Data
```{r echo=TRUE}
pa19 <- read.csv("fy19fullpropassess.csv")

# Get only desired columns
cols <- c("PID", "ZIPCODE", "PTYPE", "OWN_OCC", "LIVING_AREA", "AV_TOTAL", "YR_BUILT", "R_OVRALL_CND", "R_BDRMS", "R_FULL_BTH")

pa19 <- pa19[, cols]

# Get only "SINGLE FAM DWELLING"
pa19 <- pa19[pa19$PTYPE==101,]

# Convert string "$ddd,ddd.dd" to numeric
pa19$AV_TOTAL <- as.numeric(gsub('[$,]', '', pa19$AV_TOTAL))

summary(pa19)
```

# Analyzing the data

## Analysis
- Do the analysis as in Module3 for at least one categorical variable and at least one numerical variable. Show appropriate plots for your data.
- Do the analysis as in Module3 for at least one set of two or more variables. Show appropriate plots for your data.

## Distribution
- Pick one variable with numerical data and examine the distribution of the data.

## Central Limit Theorem
- Draw various random samples of the data and show the applicability of the
Central Limit Theorem for this variable.

## Sampling
- Show how various sampling methods can be used on your data. What are your
conclusions if these samples are used instead of the whole dataset.
- Implementation of any feature(s) not mentioned in the above specification.